Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-30 manuscript/manuscript.Rmd
mini intro messups with copying from google docs

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{bookmark}
  - \usepackage{float}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage{subcaption}
  - \usepackage{tikz-cd}
  - \usepackage{tcolorbox}
  - \usepackage{etoolbox}
  - \linenumbers
  - \doublespacing

@@

Recent advances in machine learning may allow us to now once again bridge these approaches, while also bringing new challenges of their own.
Novel data-driven methods have allowed these models to evolve into ever more complex and realistic simulations used in fisheries management, where models with over 100 parameters are not uncommon [@RamLegacyDB].
Constrained by computational limits, MDP approaches have been intractable on suitably realistic models and largely confined to more academic applications [@Costello2016].
However, advances _Deep Reinforcement Learning_, (DRL) a sub-field of machine learning, have recently demonstrated remarkable performance in a range of such MDP problems, from video games [@atari] to fusion reactions [@fusion] to the remarkable dialog abilities of ChatGPT [@chatgpt].
RL methods also bring many challenges of their own: being notoriously difficult to train and evaluate, immense computational costs, and frequent challenges with transfer-ability or reproducibility.
A review of all these issues is beyond our scope but can be found elsewhere [@rl_intro; @millie_AM-RL].
Here, though, we will focus on the issue of opacity and interpretability raised by these methods.
In contrast with optimization algorithms currently used in either ecosystem management or resource economics, RL algorithms have no guarantees of or metrics for convergence to an optimal solution.
In general, one can only assess the performance of these black box methods relative to alternatives.

<!--NEED some kind of map paragraph here saying what's ahead, what we show  -->
In this article, we We compare the performance of a DRL solution to that of a management strategy evaluation across potential 'constant mortality' policies.\footnote{A repository with all the relevant code to reproduce our results may be found at [@our-repo] in the ‘‘code’’ subdirectory. The data used is found in the ‘‘data’’ subdirectory, but the user may use the code provided to generate new data sets.} 
We consider the problem of devising harvest strategies in a three species ecosystem (Figure 1B).
We show that, somewhat surprisingly, the DRL policy closely resembles a *bang-bang* solution (see Text Box 1).
Bang-bang solutions are known to be optimal for a variety of single-species models, however, there is no a-priori reason why they should be optimal for higher-dimensional systems.
The DRL policy drives an oscillation process in the system and, in this way, maintains the long-term sustainability of fishing.
In contrast, constant mortality policies overfish at low populations and underfish (relative to the DRL solution) at high populations.
This leads to a tradeoff when leveraging constant mortality strategies: higher constant mortality rates lead to high profitability at the expense of a higher probability of overfishing to extinction, while lower constant mortality rates attain stability at the expense of profit.

@@

Throughout most of this paper we will fix the parameters of the model according to eq. \eqref{eq:param vals}.
At these values, the fixed points of the model are:\footnote{
We have included only three decimal digits for readability.
These fixed points may be generated by running the ‘‘stable\_points.py’’ script in [@repo].
}
\begin{align*}
  (X,Y,Z) \approx &(0.839, 0.054, 0.377), \quad 
    (0.048, 0.774, 0.480), \quad
    (0.359, 0.491, 0.509), \\
    &(0.666, 0.666, 0.000),\quad
    (0.899, 0.000, 0.303), \quad
    (0.000, 0.818, 0.457), \\
    &(1.000,0.000,0.000),\quad (0.000,1.000,0.000),\quad (0.000,0.000,0.000).
\end{align*}
The classification of fixed points for multi-dimensional systems is generally more complicated than in one dimension.
A fixed point might be stable against perturbations in one direction, but unstable against perturbations along another.
While a detailed classification of these fixed points is beyond the scope of this work, we point out that the point $(0.839, 0.054, 0.377)$ is stable against perturbations in all directions.
We will thus use it as a starting point for our simulations.

@@

We used the Ray framework [@ray_package] for training, in particular we used a Ray PPOConfig object to build the algorithm using the default values for all hyper-parameters (cf. [@repo] for the code used).
The agent was trained for 200 iterations of the PPO optimization step.

>>>>

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{bookmark}
  - \usepackage{float}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage{subcaption}
  - \usepackage{tikz-cd}
  - \usepackage{tcolorbox}
  - \usepackage{etoolbox}
  - \usepackage{url}
  - \linenumbers
  - \doublespacing

@@

Recent advances in machine learning may allow us to once again bridge these approaches, while also bringing new challenges of their own.
Novel data-driven methods have allowed these models to evolve into ever more complex and realistic simulations used in fisheries management, where models with over 100 parameters are not uncommon [@RamLegacyDB].
Constrained by computational limits, MDP approaches have been intractable on suitably realistic models and largely confined to more academic applications [@Costello2016].
However, advances _Deep Reinforcement Learning_, (DRL) a sub-field of machine learning, have recently demonstrated remarkable performance in a range of such MDP problems, from video games [@atari] to fusion reactions [@fusion] to the remarkable dialog abilities of ChatGPT [@chatgpt].
RL methods also bring many challenges of their own: being notoriously difficult to train and evaluate, immense computational costs, and frequent challenges with transfer-ability or reproducibility.
A review of all these issues is beyond our scope but can be found elsewhere [@rl_intro; @millie_AM-RL].
Here, though, we will focus on the issue of opacity and interpretability raised by these methods.
In contrast with optimization algorithms currently used in either ecosystem management or resource economics, RL algorithms have no guarantees of or metrics for convergence to an optimal solution.
In general, one can only assess the performance of these black box methods relative to alternatives.

<!--NEED some kind of map paragraph here saying what's ahead, what we show  -->
In this article, we compare the performance of a DRL solution to that of a management strategy evaluation across potential 'constant mortality' policies.\footnote{A repository with all the relevant code to reproduce our results may be found at \url{https://github.com/boettiger-lab/approx-model-or-approx-soln} in the ‘‘code’’ subdirectory. The data used is found in the ‘‘data’’ subdirectory, but the user may use the code provided to generate new data sets.} 
We consider the problem of devising harvest strategies in a three species ecosystem (Figure 1B).
We show that, somewhat surprisingly, the DRL policy closely resembles a *bang-bang* solution (see Text Box 1).
Bang-bang solutions are known to be optimal for a variety of single-species models, however, there is no a-priori reason why they should be optimal for higher-dimensional systems.
The DRL policy drives an oscillation process in the system and, in this way, maintains the long-term sustainability of fishing.
In contrast, constant mortality policies overfish at low populations and underfish (relative to the DRL solution) at high populations.
This leads to a tradeoff when leveraging constant mortality strategies: higher constant mortality rates lead to high profitability at the expense of a higher probability of overfishing to extinction, while lower constant mortality rates attain stability at the expense of profit.

@@

Throughout most of this paper we will fix the parameters of the model according to eq. \eqref{eq:param vals}.
At these values, the fixed points of the model are:\footnote{
We have included only three decimal digits for readability.
These fixed points may be generated by running the ‘‘stable\_points.py’’ script in \url{https://github.com/boettiger-lab/approx-model-or-approx-soln}.
}
\begin{align*}
  (X,Y,Z) \approx &(0.839, 0.054, 0.377), \quad 
    (0.048, 0.774, 0.480), \quad
    (0.359, 0.491, 0.509), \\
    &(0.666, 0.666, 0.000),\quad
    (0.899, 0.000, 0.303), \quad
    (0.000, 0.818, 0.457), \\
    &(1.000,0.000,0.000),\quad (0.000,1.000,0.000),\quad (0.000,0.000,0.000).
\end{align*}
The classification of fixed points for multi-dimensional systems is generally more complicated than in one dimension.
A fixed point might be stable against perturbations in one direction, but unstable against perturbations along another.
While a detailed classification of these fixed points is beyond the scope of this work, we point out that the point $(0.839, 0.054, 0.377)$ is stable against perturbations in all directions.
We will thus use it as a starting point for our simulations.

@@

We used the Ray framework\footnote{\url{https://docs.ray.io/}} for training, in particular we used a Ray PPOConfig object to build the algorithm using the default values for all hyper-parameters.
The agent was trained for 200 iterations of the PPO optimization step.
